---
title: "Economic and Market Monitor"
author: "Red Tortoise, LLC"
date: "May 12, 2016"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[CO,CE]{Economic and Market Monitor}
- \fancyfoot[CO,CE]{www.redtortoisellc.com}
- \fancyfoot[LE,RO]{\thepage}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE)
```
```{r echo=FALSE, warning=FALSE, message=FALSE}

library(ggplot2)
library(xts)
library(tis)
library(Quandl)
Quandl.api_key(Sys.getenv("QUANDL_TOKEN"))
chart.master<-read.csv("EconAndMktMonitor.csv",header=TRUE,stringsAsFactors = FALSE)

#' Produces standard chart. 

#' The most frequent template for a chart in this package. It is contemplated that not all charts will fit into this template and custom charts will be used. 

#' param qcode Quandl code used in Quandl function.  
#' param type Type of series returned by Quandl function, passed to Quandl. 
#' param transform Transform of data. Passed to Quandl function.  'annpct' is an exception. Calculates an
#'  annual percent change.
#' param collaps Collapse argument passed to Quandl function
#' param logscale Logscale arguement use in ggplot2
#' param main, xlab, ylab Labels passed to ggplot
#' param recession_shading True produces rectangular rectangles for recessions
#' param hline Horizontal line. Specify numeric value or 'mean' for the average
#' param trend Trend line. May be 'log' or 'linear'
std_chart<-function(qcode, type="raw", transform= NULL, collapse=NULL, logscale="",
                    main="", xlab="", ylab="", recession_shading=TRUE, hline="",
                    trend=""){
    annpct<-FALSE
    if (missing(transform)) {transform<-NULL}
    if (!is.null(transform)){
      if(transform=="annpct") {
          annpct<-TRUE
          transform<-NULL}  
    } 
    qdata<-Quandl(code=qcode,transform=transform,collapse=collapse)
    data.freq<-attr(qdata,"freq")
    obsperyr<-switch(data.freq,
                 daily=365,
                 weekly=52,
                 quarterly=4,
                 monthly=12,
                 annual=1)
    if (nrow(qdata)==0){
        stop(paste("No rows returned in call to Quandl with code",qcode))
    }
    data<-qdata
    if ("Last" %in% colnames(qdata)) {
        data<-data.frame(DATE=qdata$Date,VALUE=qdata$Last)
    }
    if ("VALUE" %in% colnames(qdata)) {
        data<-data.frame(DATE=qdata$DATE,VALUE=qdata$VALUE)
    }
    if ("Value" %in% colnames(qdata)) {
        data<-data.frame(DATE=qdata$Date,VALUE=qdata$Value)
    }
    if ("Adjusted Close" %in% colnames(qdata)) {
        data<-data.frame(DATE=qdata$Date,VALUE=qdata[,"Adjusted Close"])
    }
    if ("VIX Close" %in% colnames(qdata)) {
        data<-data.frame(DATE=qdata$Date,VALUE=qdata[,"VIX Close"])
    }
    if ("Settle" %in% colnames(qdata)) {
        data<-data.frame(DATE=qdata$Date,VALUE=qdata[,"Settle"])
    }
    if (nrow(data)==0){
        stop(paste("No rows when Quandl coverted to data with code",qcode))
    }
    if (annpct){
        data<-data.frame(
            DATE=data$DATE[1:(nrow(data)-obsperyr)],
            VALUE=rollapply(data$VALUE,obsperyr+1,function (x) 100*x[1]/x[obsperyr+1]-100))
    }
    if (trend=="log"){
        data.trend<-data.frame(y=log(data$VALUE),x=as.numeric(data$DATE))
        new<-data.frame(x=data.trend$x)
        expo.mdl<-lm(y~x,data=data.trend)
        expo.pred<-predict(expo.mdl,new)
        data.trend<-data.frame(DATE=data$DATE,TREND=exp(expo.pred))
    }
    if (trend=="linear"){
        data.trend<-data.frame(y=data$VALUE,x=as.numeric(data$DATE))
        new<-data.frame(x=data.trend$x)
        expo.mdl<-lm(y~x,data=data.trend)
        expo.pred<-predict(expo.mdl,new)
        data.trend<-data.frame(DATE=data$DATE,TREND=expo.pred)
    }
    if(recession_shading){
            recessions<-nberDates()
            temp<-as.Date(as.character(recessions),"%Y%m%d")
            recessions<-data.frame(Start=temp[seq(1,nrow(recessions))],
                                   End=temp[(nrow(recessions)+1):(nrow(recessions)*2)])
            recessions.trim<-subset(recessions,End>=min(data$DATE))
            if (nrow(recessions.trim)<1) {
                recession.shading<-FALSE # no recessions to shade
            } else {
                if (recessions.trim$Start[1]<min(data$DATE)) {recessions.trim$Start[1]<-min(data$DATE)}
            }
            
    }
    out<-ggplot(data=data)+geom_line(aes(x=DATE,y=VALUE),color="red")+ 
        labs(title=main,x=xlab,y=ylab)+
        scale_fill_manual(values=cbPalette)
    if (trend!=""){
        out<-out+geom_line(aes(x=DATE,y=TREND),color="darkgray",data=data.trend)
    }
    out<- out + theme_bw()+theme(plot.title=element_text(size=10, face="bold"),axis.title=element_text(size=10))
    if (logscale!="") {
        out<-out + scale_y_continuous(trans = logscale)
    }    
    if (recession_shading) {
        out<-out + geom_rect(data=recessions.trim,
                   aes(xmin=Start,xmax=End,ymin=min(data$VALUE),ymax=Inf), fill="darkgray",alpha=0.2)}
    if (hline!=""){
        if (is.numeric(hline)){
            out<-out+geom_hline(yintercept=hline, col="lightblue")    
        } else {
            temp<-do.call(hline,list(data$VALUE))
            out<-out+geom_hline(yintercept=temp, col="lightblue")
        }
    }
    
    return(out)
}

```

```{r echo=FALSE, fig.width=3.5, fig.height=3.5, warning=FALSE, comment=""}
par(mfrow=c(1,2))
for (i in 1:nrow(chart.master)){
    if (is.na(chart.master[i,"transform"])){
        xform<-NULL    
    } else {xform<-chart.master[i,"transform"]}
    if (is.na(chart.master[i,"collapse"])){
        clapse<-NULL    
    } else {clapse<-chart.master[i,"collapse"]}
    if (is.na(chart.master[i,"xlab"])){
        xlabel<-""
    } else {xlabel<-chart.master[i,"xlab"]}
    if (is.na(as.numeric(chart.master[i,"hline"]))){
        horizline<-chart.master[i,"hline"]
    } else {horizline<-as.numeric(chart.master[i,"hline"])}
    cht<-std_chart(qcode=chart.master[i,"qcode"],
              type=chart.master[i,"type"],
              transform= xform,
              collapse = clapse,
              logscale = chart.master[i,"logscale"],
              main=chart.master[i,"main"],
              xlab=xlabel,
              ylab=chart.master[i,"ylab"],
              recession_shading = chart.master[i,"recession_shading"],
              hline = horizline, 
              trend=chart.master[i,"trend"])
    print(cht)
}

```
   
Source: [Quandl](www.quandl.com) (codes include: FRED/GDPC96, FRED/CPIAUCSL, FRED/CPILFESL, FRED/UNRATE, FRED/ALTSALES, FRED/MEHOINUSA672N, FRED/DSPI, FRED/HOUST, FRED/T10YIE, FRED/CP, FRED/STLFSI, CHRIS/CME_CL1, CHRIS/ICE_B1, MULTPL/SHILLER_PE_RATIO_MONTH, FRED/DGS10, FRED/DGS2, FRED/DTB3, CBOE/VIX, FRED/DTWEXM, FRED/DTWEXB, YAHOO/INDEX_GSPC)

Red Tortoise does not guaranty or make any warranties about the accuracy or completeness of this data. Some data may be inaccurate, incomplete or unfit for a particular purpose.  This document does not contain investment advice.